# Welcome

```{r include = FALSE}
source("_common.R")
```

It's a vast world out there. We're going to cover a lot of ground. Don't be scared. It's going to be **so. much. FUN!**

I'm just playing right now and specifically thinking about code styling. So here's a code chunk for producing a very basic plot.

```{r}
# load the libraries
library(tidyverse)

# create a plot
ggplot(mpg, aes(displ, cty)) +
  geom_point(color = "#B3E5D9") +
  geom_smooth(color = "#7EC1E7",
              fill = "#C7E4F5") +
  theme_minimal(15)
```

We could of course create some summary data too.

```{r}
# create some summary data
mpg %>% 
  group_by(cyl) %>% 
  summarize(mean_cty = mean(cty),
            sd_cty = sd(cty))
```

But the above doesn't give us an example of a string, which we'll also use a lot, so let's try an example with that.

```{r}
mpg %>% 
  filter(manufacturer == "audi")
```

Finally, let's write a quick function that has a logical in it (since we haven't tested those yet).

```{r}
mn <- function(x) {
  mean(x, na.rm = TRUE)
}
mn(mpg$hwy)
```

And overall, I'm pretty happy with how that's all looking.